---
title: "Tutorial for RNASeq analysis using pathviewwrap package"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Tutorial for RNASeq analysis using pathviewwrap package}
  %\VignetteEngine{knitr::knitr} 
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>"
)
```

```{r setup}
library(pathviewwrap)
```

Pathviewwrap is a wrapper function/package that allows users to analyse the RNASeq data from the raw files to the pathway visualization.
It does quality control analysis of raw files, performs adapter and quality trimming, builds genome index and does alignemnet , counts genes
perfoms differential gene expression analysis, does gene enrichment test using GAGE and visualize the enriched pathways using pathview all using one wrapper function.
It has the ability to  continue the analysis if it is halted at any stage and generate quality pictures and generate comprehensive analysis of the data.

```{r }
data(bods)
```
# Session info

```{r sessionInfo}
sessionInfo()
``` 


