---
title: "Latest Tutorial for RNASeq analysis using pathviewwrap package"
author: "Eliza Dhungel"
package: pathviewwrap
output: 
    rmarkdown::html_document:
    highlight : pygments
    toc : true
    fig_width: 5
vignette: >
    %\VignetteIndexEntry{Latest Tutorial for RNASeq analysis using pathviewwrap }
    %\VignetteEngine{knitr::knitr} 
    %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>"
)
```

## Installation

To install pathviewwrap package, start R(version "4.3") and enter:

```{r setup, eval = FALSE}

if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")
BiocManager::install(pathviewwrap)
```

Based on the species you want to analyse, you can install the bioconductor 
annotation and genome package. For example, for human species

```{r installation, eval = FALSE}
library(pathviewwrap)
data(anntpkglist, package = "pathviewwrap")
genomepkg <- anntpkglist$gneome[which(anntpkglist$species=="Homo sapiens")]
anntpkg <- anntpkglist$gneome[which(anntpkglist$species=="Homo sapiens")]
BiocManager::install(genomepkg)
BiocManager::install(anntpkg)
```

## Documentation
browseVignettes("pathviewwrap")

## Usage
```{r wrapper, eval = FALSE}
library(pathviewwrap)
Results <- tempdir()
library(pathviewwrap)
system.time({
    pathviewwrap(
    ref.dir = NA,
    phenofile = system.file(
    package = "pathviewwrap",
    "extdata/phenotest_ungz.txt"
    ),
outdir = Results, entity = "Mus musculus", corenum = 16,
    compare = "as.group", seq_tech = "Illumina", keep_tmp = TRUE,
    rerun = FALSE, diff.tool = "DESeq2", aligner = "Rhisat2"
    )
})
```


## Overview
Pathviewwrap is a wrapper function/package that allows users to analyse the 
RNASeq data from the raw files to the pathway visualization.
It does quality control analysis of raw files, performs adapter and quality 
trimming, builds genome index and does alignemnet , counts genes perfoms 
differential gene expression analysis, does gene enrichment test using GAGE and
visualize the enriched pathways using pathview all using one wrapper function.
It has the ability to  continue the analysis if it is halted at any stage and 
generate quality pictures and generate comprehensive analysis of the data.

With one wrapper function, it runs all the steps listed below.

**STEP 1 : Quality control**

**STEP 1a: running fastqc**

It runs fastqc analysis in R using fastqcr. If fastqc is not available in system
to run by R, this function is capable of downloading the fastqc tools before 
running the quality check. The results are standard html files where the 
quality of each fastq files can be examined.

**STEP 1b : running fastp**

After running fastqc, it runs fastp. The function takes name of the samples and
for each sample does the quality and adapter trimming for Illumina and long read
sequencing. It works for both PE and SE data. HTML files are generated for each
fastq files that has information/figures of quality control before and after 
quality trimming.

**STEP 2: making txdb obj**

The wrapper then makes TxDb object either from annnotation file or by loading 
from the annotation package. Make_txdbobj uses the makeTxDbFromGFF to make TxDb 
object from transcript annotations available in gtf file in ref.dir or if 
ref.dir is NA, it makes txdb object fromt the annotaion package

**STEP 3 : Alignmnet and counting**

After the txdb object is formed, the wrapper runs the Rhisat2 or Rbowtie for 
alignment on paired or single end mode depending on data. It saves the alignment
object in RDS file which can be loaded in R for further analysis. If the 
reference index is not found in the reference directory, it creates reference 
index before running alignment. If the refernece genome is a package the 
reference index is created as R package. It generates the barplot of mapped 
and unmapped sequence reads.

**STEP 4: counting aligned sequences**

After aligning the reads to reference genome, the wrapper generates the count of
gene and store it in a table with genes in row and counts in columns. The table 
is stored as a RDS file that can be loaded into R for future analysis and if 
gene id is ensembl, the wrapper converts it to entrez to match with genes names 
for gene set analysis.

**STEP 5a ; running differential gene analysis using DESeq2**

Then the wrapper runs standard DESeq2 for differential gene expression analysis
and plots volcano plots. The function run_deseq2 takes counts and the list 
indicating reference and samples and the directory where the results are stored 
and performs the deseq2 analysis. The output is result table with columns of 
genes and log2FoldChange from result of deseq2 analysis and a volcanoplot.

**STEP 5b ; running differential gene analysis using edgeR**

Then the wrapper runs standard edgeR for differential gene expression analysis 
and plots volcano plots. The function run_deseq2 takes counts and the list 
indicating reference and samples and the directory where the results are stored 
and performs the deseq2 analysis. The output is result table with columns of 
genes and log2FoldChange from result of deseq2 analysis and a volcanoplot.

**STEP 6 : running pathway analysis using GAGE**

After the differential gene analysis the wrapper runs generally applicable gene 
set enrichment for pathway analysis, GAGE based upon the user supplied 
comparison method for the species specified. The biological process, cellular 
component and molecular function analysis for GO terms are done separately. 
Also, KEGG disease and KEGG signalling and metabolism pathways are analysed 
separately.

**STEP 7: visualizing the pathway using Pathview**

Finally the top enriched pathways with "q.val" < 0.1 are visualized using 
pathview.


## Acknowledgements


## References

* Luo W, Brouwer C. Pathview: an R/Biocondutor package for pathway-based data 
integration and visualization. Bioinformatics, 2013, 29(14):1830-1831, 
doi: 10.1093/bioinformatics/btt285

* Luo, W., Friedman, M.S., Shedden, K. et al. GAGE: generally applicable gene 
set enrichment for pathway analysis. BMC Bioinformatics 10, 161 (2009).
https://doi.org/10.1186/1471-2105-10-161





# Session info
```{r sessionInfo}
sessionInfo()
``` 


